---
title: "30708974_Assignment1"
author: "Malav Parikh"
date: "20/03/2021"
output: html_document
---

## Task A Investigating Natural Increase in Australia's population


## A1. Investigating the Births, Deaths and TFR Data

## 1. Number of births recorded in each state/territory for
## different Australian states over the years
```{r}
# loading libraries
library(tidyverse)
library(ggplot2)
# reading csv
births = read_csv('FIT5145 A1 Data/Task A/Births.csv')
# checking head of data
head(births)

# converting to long form
birth_long = births %>% gather(key = State, value = Count, -Year)
head(birth_long)
```
```{r}
## 1(A) Plotting number of births recorded in each state over the years.
birth_long %>% ggplot(aes(x = Year, y = Count, colour = State)) +
  geom_point() +
  facet_wrap(~State) +
  theme(axis.text.x = element_text(angle = 60)) +
  ggtitle('Number of births in each state in each year')
```
```{r}
## 1(B)
birth_long %>% filter(Year == 2016) %>% 
  ggplot(aes(x = reorder(State,Count), y = Count)) +
  geom_bar(stat="identity", fill = "steelblue") +
  xlab('State') +
  ggtitle('Number of births in each Australian state in 2016')
  
```
## 2. Total Fertility Rate for Australian states by years
```{r}
# reading data from csv file
tfr = read_csv('FIT5145 A1 Data/Task A/TFR.csv')
# checking head of data
head(tfr)
# converting to long format
tfr_long = tfr %>% gather(key = State, value = Rate, -Year)
# filtering data to get data for QLD and NT only
tfr_long_filter = tfr_long %>% filter(State %in% c('QLD','NT'))
# checking head of data
head(tfr_long_filter)
```


```{r}
## 2(a)
## Minimum value of TFR in Queensland with year

# filtering fertility rate data for Queensland
QLD_tfr_min = tfr_long_filter %>% filter(State == 'QLD', Rate == min(Rate))
QLD_tfr_min
```


```{r}
## 2(b)
## Value of TFR in NT for the year with lowest tfr in Queensland.

# filtering fertility rate data for NT for a specific year
tfr_long_filter %>% filter(State == 'NT', Year == QLD_tfr_min$Year)
```
## 3. Natural Growth in  Australia's Population over years
```{r}
# loading death data
deaths = read_csv('FIT5145 A1 Data/Task A/Deaths.csv')
# checking head of data
head(deaths)
# converting to long format
deaths_long = deaths %>% gather(key = State, value = Count, -Year)
# checking head of data 
head(deaths_long)
```


```{r}
# Plotting the natural growth in AUstralia's population over the years
# grouping births by year
birth_by_year = birth_long %>%group_by(Year)%>%summarise(total_births = sum(Count))
# grouping deaths by year
death_by_year = deaths_long%>%group_by(Year)%>%summarise(total_deaths = sum(Count))

# preparation of natural growth data
Natural_growth = merge(birth_by_year,death_by_year, by = 'Year')
Natural_growth$natural_growth_rate = Natural_growth$total_births - Natural_growth$total_deaths
# checking head of data
head(Natural_growth)

# plotting natural growth rate trend
ggplot(Natural_growth, aes(x=Year,y=natural_growth_rate)) +
  geom_line(color='yellowGreen') +
  ylab('Natural growth rate') +
  ggtitle('Natural growth in Australia over years')
```
## A2. Investigating the Migration Data

## 1. Net overseas Migration Data
```{r}
## 1(A)
# reading data from csv
nom = read_csv('FIT5145 A1 Data/Task A/NOM.csv')
# checking head of data
head(nom)
# converting data to long form
nom_long = nom %>% gather(key = State, value = nom_rate, -Year)
# checking head of data
head(nom_long)

# plotting NOM data for 3 states
ggplot(nom_long %>% filter(State %in% c("VIC", "TAS", "WA"))) +
 aes(x = Year, y = nom_rate, colour = State) +
 geom_point(size = 1L) +
 geom_smooth() +
 scale_color_hue() +
 facet_grid(vars(), vars(State)) +
 theme(axis.text.x = element_text(angle = 60)) +
 ylab('Net Overseas Migration') +
  ggtitle('Net overseas Migration over years')
```


```{r}
## 1(B) Net Overseas Migration over time in Australia
# grouping data by year to get the total overseas migration data for each year
nom_yearly = nom_long %>% group_by(Year) %>% summarise(net_overseas_migration = sum(nom_rate))
#checking head of data
head(nom_yearly)

# plotting Net overseas migration over time in Australia
ggplot(nom_yearly, aes(x=Year,y=net_overseas_migration)) +
  geom_line(color = 'violetred2') +
  scale_y_continuous(labels = scales::comma) +
  ylab('Net Overseas Migration') +
  ggtitle('Net Overseas Migration to Australia')
```
## 2. Relation between Net overseas Migration and Net Interstate Migration
```{r}
# 2(A) 
# reading data from csv
nim = read_csv('FIT5145 A1 Data/Task A/NIM.csv')
# checking head of data
head(nim)
# converting to long form
nim_long = nim %>% gather(key = State, value = nim_rate, -Year)
# checking head of data
head(nim_long)
# merging nim and nom data into one
combined_NOM_NIM = merge(nom_long,nim_long, by = c('Year','State'))
# checking head of data
head(combined_NOM_NIM)

# first year and last year in data
first_year = first(combined_NOM_NIM$Year)
last_year = last(combined_NOM_NIM$Year)

first_year
last_year
```
```{r}
#2(B) Relation between NIM and NOM

# plotting graph for nim vs nom
ggplot(combined_NOM_NIM, aes(x = nom_rate, y = nim_rate)) +
    geom_point(alpha=0.5, size=2, color = 'slateblue') +
    xlab('Net Overseas Migration') +
    ylab('Net Interstate Migration') +
  ggtitle('NOM vs NIM')
```

```{r}
## 2(C) NIM and NOM for VIC
# plotting scatter plot for NIM vs NOM in Victoria
ggplot(combined_NOM_NIM %>% filter(State == 'VIC')) +
      aes(x = nom_rate, y = nim_rate)+
      geom_point(alpha=0.5, size=2, color = 'tomato') +
      xlab('Net Overseas Migration') +
      ylab('Net Interstate Migration') +
      ggtitle('NOM vs NIM in Victoria')
```


```{r}
## 2(D) NIM for QLD and NSW
# plotting NIM for NSW and QLD
ggplot(nim_long %>% filter(State %in% c('QLD','NSW'))) +
      aes(x = Year, y = nim_rate, colour = State)+
      geom_point() +
      geom_smooth() +
      ylab('Net Interstate Migration') +
      ggtitle('NIM over years in NSW and QLD')
```

## Task B Exploratory Data Analysis about Bushfires in Australia


## B1. Investigating the twitter data

##Twitter data analysis
```{r}
# reading data from csv
twitter_data = read_csv('FIT5145 A1 Data/Task B/twitter_data.csv')
# checking head of data
head(twitter_data)

```


```{r}
## 1 
#Total number of tweets 
Total_tweets = sum(twitter_data$`#tweetsPosted`)
Total_tweets

#Tweets from verified account
Verified_tweets = twitter_data %>% filter(isVerified == 1) %>% group_by(isVerified) %>% summarise(total = sum(`#tweetsPosted`))
Verified_tweets
```


```{r}
## 2 Plotting Distribution of entities 
ggplot(twitter_data) +
 aes(x = `#entities`) +
 geom_histogram(bins = 11,binwidth = 1,color = 'black', fill = 'yellow') +
  xlab('Number of entities') +
  ggtitle('Distribution of number of entities')
```
```{r}
## 3 Descriptive statistics for relevant tweets
Relevant_tweets = twitter_data %>% filter(relevanceJudge == 1)

mean_value = mean(Relevant_tweets$'#entities')
std_1 = sd(Relevant_tweets$'#entities')
quartile_1 = quantile(Relevant_tweets$'#entities', 0.25)
median_value = median(Relevant_tweets$'#entities')
maximum_value = max(Relevant_tweets$'#entities')

mean_value
std_1
quartile_1
median_value
maximum_value

```
```{r}
## 3 Descriptive statistics for non-relevant tweets
Non_Relevant_tweets = twitter_data %>% filter(relevanceJudge == 0)

mean_value = mean(Non_Relevant_tweets$'#entities')
std = sd(Non_Relevant_tweets$'#entities')
quartile_1 = quantile(Non_Relevant_tweets$'#entities', 0.25)
median_value = median(Non_Relevant_tweets$'#entities')
maximum_value = max(Non_Relevant_tweets$'#entities')

mean_value
std
quartile_1
median_value
maximum_value
```

```{r}
## 4 Average length of tweets by relevance of tweet
# grouping data by relevance judge and getting average length of tweet
Average_length_of_tweets = twitter_data %>% group_by(relevanceJudge) %>% summarise(average_length = round(mean(length),2))  
# printing the results
Average_length_of_tweets
```

```{r}
## 5 Further data analysis
head(twitter_data)

# adding age category column and updating its value based on age(time since author joined twitter)
twitter_data$AgeCategory <-
  ifelse((twitter_data$twitterAge >= 0 & twitter_data$twitterAge < 1), '0-1',
         ifelse((twitter_data$twitterAge >= 1 & twitter_data$twitterAge < 2), '1-2', 
                ifelse((twitter_data$twitterAge >= 2 & twitter_data$twitterAge < 3), '2-3', 
                       ifelse((twitter_data$twitterAge >= 3 & twitter_data$twitterAge < 4), '3-4',
                              ifelse((twitter_data$twitterAge >= 4 & twitter_data$twitterAge < 5), '4-5','5+')))))

head(twitter_data)
```

```{r}
## 5(A) box plot for each age group
box_plot <- ggplot(twitter_data, aes(x=AgeCategory, y=length, fill=AgeCategory)) + 
    geom_boxplot() +
    xlab('Age Category') +
    ylab('Tweet Length')
box_plot
```

```{r}
## 5(B) Lowest and highest Median tweet length 
twitter_data %>% group_by(AgeCategory) %>% summarise(median_tweet_length = median(length))
```

```{r}
## 5(C) Age group activity
Activity = twitter_data %>% group_by(AgeCategory) %>% summarise(total_tweets = sum(`#tweetsPosted`))
Activity
```

```{r}
## 5(D) Total tweets by age group
ggplot(Activity,aes(x=reorder(AgeCategory, total_tweets),y=total_tweets))+
  geom_bar(stat = 'identity', fill = 'paleturquoise') +
  xlab('Age Category') +
  ylab('Total Number of tweets') +
  ggtitle('Number of tweets by age category')
```

```{r}
## 5(E) Followers by age group
No_of_followers = twitter_data %>% group_by(AgeCategory) %>% summarise(total_followers = sum(nFollowers))

No_of_followers
```